Model Search in Relational Domains

### 2016-10-22 algorithm.tex
comments in sec4

\section{Abstract Conflict Driven Clause Learning}
%
\begin{algorithm2e}[t]
\DontPrintSemicolon
\SetKw{return}{return}
\SetKwRepeat{Do}{do}{while}
%\SetKwFunction{assume}{assume}
%\SetKwFunction{isf}{isFeasible}
\SetKwData{conflict}{conflict}
\SetKwData{safe}{safe}
\SetKwData{sat}{sat}
\SetKwData{unsafe}{unsafe}
\SetKwData{unknown}{unknown}
\SetKwData{true}{true}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKwFor{Loop}{Loop}{}{}
\SetKw{KwNot}{not}
\begin{small}
\Input{A program in the form of a set of abstract transformers $\abstransset$.}
\Output{The status \safe or \unsafe, and a counterexample if \unsafe.}
$\trail \leftarrow \langle\rangle$ \;
$\reasons \leftarrow []$ \;
$\mathit{result} \leftarrow \deduce_{\propheur}(\abstransset,\trail,\reasons)$ \;
\lIf{$\mathit{result}$ = \conflict} {
  \return \safe}
\While{$true$} 
{
\lIf{$\mathit{result}$ = \sat} {
  \return (\unsafe,$\abs(\trail)$)}
  $\decisionvar \leftarrow \decide_{\decheur}(\abs(\trail))$ \;
  $\trail \leftarrow \trail \cdot \decisionvar$ \; 
  $\reasons[|\trail|] \leftarrow \top$ \;
  $\mathit{result} \leftarrow \deduce_{\propheur}(\abstransset,\trail,\reasons)$\;
  \Do{$\mathit{result} = \conflict$} {
    \lIf{$\neg \analyzeconflict_{\confheur}(\abstransset,\trail,\reasons)$} {
      \return \safe
    }
    $\mathit{result} \leftarrow \deduce_{\propheur}(\abstransset,\trail,\reasons)$ \;
  }
}
\end{small}
\caption{Abstract Conflict Driven Clause Learning $ACDCL_{\propheur,\decheur,\confheur}(\abstransset)$ \label{Alg:acdcl}}
\end{algorithm2e}
%
%\subsection{Algorithm for Abstract Conflict Driven Clause Learning}
Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
The algorithm takes a program \pscmt{make this a set of transformers
  (parametrised by the propagation heuristics)}
$\abstransset$ and propagation, decision, and conflict analysis
heuristics as inputs.  
\rmcmt{Approximation of the concrete transformers in 
$\abstransset$ are typically available in abstract domain in the 
form of strongest-post condition or weakest pre-condition. } 
As internal state, it maintains a propagation
trail $\trail$ and a reason trail \pscmt{map?} $\reasons$.
The propagation trail is an initially empty sequence of meet
irreducibles inferred by the abstract model search phase (deductions
and decisions). The reason trail maps the elements of the
propagation trail to transformers $s\in\abstransset$ that were used to
derive them. The \emph{abstract value} $\abs(\trail)$
corresponding to the propagation trail $\trail$ is the
conjunction of the meet irreducibles on the trail:
$\abs(\trail)=\bigsqcap_{m \in \trail}m$ with
$\abs(\trail)=\top$ if $\trail$ is the empty
sequence $\langle\rangle$.

The procedure $\deduce$ computes a greatest fixed point over the
transformers in $\abstransset$ that refines the abstract value,
similar to the BCP step in SAT solvers.  If the result of $\deduce$
returns \textsf{conflict} ($\bot$), the algorithm terminates with
\textsf{safe}.  Else, the analysis enters into the while($\true$) loop
and makes a new decision by a call to $\decide$ which returns a new
meet irreducible $\decisionvar$.
%
\pscmt{what happens if we cannot make any decision any more?
  -- then we should either have deduced $\bot$ or $\abs(\trail)$ must
  be $\gamma$-complete} 
%
We concatenate $\decisionvar$ to the
trail $\trail$.  The decision $\decisionvar$ therefore refines the current
abstract value $\abs(\trail)$ represented by the trail.
% WE ExPLAIN THAT LATER
%However, $\decisionvar$ must be consistent with the state of the
%solver, i.e. $\abs(\trail\cdot \decisionvar)\neq \bot$.
%
For example, a decision in the interval domain restricts the range of 
intervals for variables.
%
We set the corresponding entry in the reason trail $\reasons$ to $\top$
to mark it as a decision.
% so the analysis jumps under a 
%greatest fixed-point \pscmt{???}. Note that the widening operation in abstract interpretation 
%jumps above a least fixed-point, so decisions can be viewed as 
%{\em dual widening} \pscmt{what do we learn?}.

@@

In the latter case the algorithm enters in the $\analyzeconflict$
phase to learn the reason for the conflict.   There can be multiple
incomparable reasons for conflict \pscmt{do you have to explain this here?} -- based on the choice of Unique
Implication Point (UIP)~\cite{cdcl}.  ACDCL heuristically chooses one
reason.  A learnt clause must include asserting cuts which guarantees
derivation of new meet irreducibles after backtracking. The clause
learning and backtracking continues as long as the result of deduction
is conflicting ($\bot$), that is, the abstract value does not
abstractly satisfy the formula \pscmt{define}.  If no further
backtrack is possible, then the algorithm terminates and $\formula$ is
\textsf{safe}. Else, the algorithm makes a new decision and the above
process is repeated until $\deduce$ returns \textsf{sat} or the
algorithm backtracks to decision level 0 after a conflict in which
case it returns \textsf{safe}.

>>>>

\section{Abstract Conflict Driven Clause Learning}
%
\begin{algorithm2e}[t]
\DontPrintSemicolon
\SetKw{return}{return}
\SetKwRepeat{Do}{do}{while}
%\SetKwFunction{assume}{assume}
%\SetKwFunction{isf}{isFeasible}
\SetKwData{conflict}{conflict}
\SetKwData{safe}{safe}
\SetKwData{sat}{sat}
\SetKwData{unsafe}{unsafe}
\SetKwData{unknown}{unknown}
\SetKwData{true}{true}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKwFor{Loop}{Loop}{}{}
\SetKw{KwNot}{not}
\begin{small}
\Input{A program in the form of a set of abstract transformers $\abstransset$.}
\Output{The status \safe or \unsafe, and a counterexample if \unsafe.}
$\trail \leftarrow \langle\rangle$ \;
$\reasons \leftarrow []$ \;
$\mathit{result} \leftarrow \deduce_{\propheur}(\abstransset,\trail,\reasons)$ \;
\lIf{$\mathit{result}$ = \conflict} {
  \return \safe}
\While{$true$} 
{
\lIf{$\mathit{result}$ = \sat} {
  \return (\unsafe,$\abs(\trail)$)}
  $\decisionvar \leftarrow \decide_{\decheur}(\abs(\trail))$ \;
  $\trail \leftarrow \trail \cdot \decisionvar$ \; 
  $\reasons[|\trail|] \leftarrow \top$ \;
  $\mathit{result} \leftarrow \deduce_{\propheur}(\abstransset,\trail,\reasons)$\;
  \Do{$\mathit{result} = \conflict$} {
    \lIf{$\neg \analyzeconflict_{\confheur}(\abstransset,\trail,\reasons)$} {
      \return \safe
    }
    $\mathit{result} \leftarrow \deduce_{\propheur}(\abstransset,\trail,\reasons)$ \;
  }
}
\end{small}
\caption{Abstract Conflict Driven Clause Learning $ACDCL_{\propheur,\decheur,\confheur}(\abstransset)$ \label{Alg:acdcl}}
\end{algorithm2e}
%
%\subsection{Algorithm for Abstract Conflict Driven Clause Learning}
Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
The input to the algorithm is a program in the form of the set of its abstract transformers $\abstransset$ w.r.t.\ an abstract domain~$\domain$. The algorithm is parametrised by propagation, decision, and conflict analysis
heuristics.  
%\rmcmt{Approximation of the concrete transformers in 
%$\abstransset$ are typically available in abstract domain in the 
%form of strongest-post condition or weakest pre-condition. } 
It maintains a propagation
trail $\trail$ and a reason trail $\reasons$.
The propagation trail is an initially empty sequence of meet
irreducibles inferred by the abstract model search phase (deductions
and decisions). The reason trail maps the elements of the
propagation trail to transformers $\abstrans\in\abstransset$ that were used to
derive them. The \emph{abstract value} $\abs(\trail)$
corresponding to the propagation trail $\trail$ is the
conjunction of the meet irreducibles on the trail:
$\abs(\trail)=\bigsqcap_{m \in \trail}m$ with
$\abs(\trail)=\top$ if $\trail$ is the empty
sequence $\langle\rangle$.

The procedure $\deduce$ computes a greatest fixed point over the
transformers in $\abstransset$ that refines the abstract value,
similar to the BCP step in SAT solvers.  If the result of $\deduce$
returns \textsf{conflict} ($\bot$), the algorithm terminates with
\textsf{safe}.  Else, the analysis enters into the while($\true$) loop
and makes a new decision by a call to $\decide$ which returns a new
meet irreducible $\decisionvar$.
%
%\pscmt{what happens if we cannot make any decision any more?
%  -- then we should either have deduced $\bot$ or $\abs(\trail)$ must
%  be $\gamma$-complete} 
%
We concatenate $\decisionvar$ to the
trail $\trail$.  The decision $\decisionvar$ therefore refines the current
abstract value $\abs(\trail)$ represented by the trail.
% WE ExPLAIN THAT LATER
%However, $\decisionvar$ must be consistent with the state of the
%solver, i.e. $\abs(\trail\cdot \decisionvar)\neq \bot$.
%
For example, a decision in the interval domain restricts the range of 
intervals for variables.
%
We set the corresponding entry in the reason trail $\reasons$ to $\top$
to mark it as a decision.
% so the analysis jumps under a 
%greatest fixed-point \pscmt{???}. Note that the widening operation in abstract interpretation 
%jumps above a least fixed-point, so decisions can be viewed as 
%{\em dual widening} \pscmt{what do we learn?}.

@@

In the latter case the algorithm enters in the $\analyzeconflict$
phase to learn the reason for the conflict.   There can be multiple
incomparable reasons for conflict.
% \pscmt{do you have to explain this here?} -- based on the choice of Unique
%Implication Point (UIP)~\cite{cdcl}.  
ACDCL heuristically chooses one reason and learns it by adding at as
an abstract transformer to $\abstransset$. Then it backtracks by
removing the tail of the $\trail$ up to a point where it becomes
consistent, and continues making deductions from there.
If $\analyzeconflict$ returns $\false$ then $\formula$ is unsatisfiable
and we return  \textsf{safe}.

% A learnt clause must include asserting cuts which guarantees
% derivation of new meet irreducibles after backtracking. The clause
% learning and backtracking continues as long as the result of deduction
% is conflicting ($\bot$), that is, the abstract value does not
% abstractly satisfy the formula \pscmt{define}.  If no further
% backtrack is possible, then the algorithm terminates and $\formula$ is
% \textsf{safe}. Else, the algorithm makes a new decision and the above
% process is repeated until $\deduce$ returns \textsf{sat} or the
% algorithm backtracks to decision level 0 after a conflict in which
% case it returns \textsf{safe}.
